---
title: 'Data integration and clustering of cardiomyocytes: fetal, young, adult'
author: "Belinda Phipson"
date: "10/17/2019"
output: html_document
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
    echo = TRUE,
    message = FALSE,
    warning = FALSE
)
```

# Load libraries and functions

```{r}
library(edgeR)
library(RColorBrewer)
library(org.Hs.eg.db)
library(limma)
library(Seurat)
library(monocle)
library(cowplot)
library(DelayedArray)
library(scran)
library(NMF)
library(workflowr)
library(ggplot2)
library(clustree)
library(dplyr)
```

```{r}
source("/misc/card2-single_cell_nuclei_rnaseq/Porello-heart-snRNAseq/code/normCounts.R")
source("/misc/card2-single_cell_nuclei_rnaseq/Porello-heart-snRNAseq/code/findModes.R")
source("/misc/card2-single_cell_nuclei_rnaseq/Porello-heart-snRNAseq/code/ggplotColors.R")
```

```{r readTargets}
targets <- read.delim("/misc/card2-single_cell_nuclei_rnaseq/Porello-heart-snRNAseq/data/targets.txt",header=TRUE, stringsAsFactors = FALSE)
targets$FileName2 <- paste(targets$FileName,"/",sep="")
targets$Group_ID2 <- gsub("LV_","",targets$Group_ID)
group <- c("Fetal_1","Fetal_2","Fetal_3",
           "Young_1","Young_2","Young_3",
           "Adult_1","Adult_2","Adult_3", 
           "Diseased_1","Diseased_2",
           "Diseased_3","Diseased_4")
m <- match(group, targets$Group_ID2)
targets <- targets[m,]
```

```{r readData}
fetal.integrated <- readRDS(file="./output/RDataObjects/fetal-int.Rds")
load(file="./output/RDataObjects/fetalObjs.Rdata")

young.integrated <- readRDS(file="./output/RDataObjects/young-int.Rds")
load(file="./output/RDataObjects/youngObjs.Rdata")

adult.integrated <- readRDS(file="./output/RDataObjects/adult-int.Rds")
load(file="./output/RDataObjects/adultObjs.Rdata")
```

# Set default clustering resolution

```{r}
# Default 0.3
Idents(fetal.integrated) <- fetal.integrated$integrated_snn_res.0.3
DimPlot(fetal.integrated, reduction = "tsne",label=TRUE,label.size = 6)+NoLegend()

# Default 0.3
DimPlot(young.integrated, reduction = "tsne",label=TRUE,label.size = 6)+NoLegend()

# Default 0.6
DimPlot(adult.integrated, reduction = "tsne",label=TRUE,label.size = 6)+NoLegend()
```

# Merge all data together

```{r}
heart <- merge(fetal.integrated, y = c(young.integrated, adult.integrated), project = "heart")
table(heart$orig.ident)
```

# Get cardiomyocytes only

```{r}
cardio <- subset(heart,subset = Broad_celltype == "Cardiomyocytes")
```

# Check for poor quality cells

Cardiomyocytes are fairly large cells and we wouldn't expect them to only be expressing very few genes.

```{r}
par(mfrow=c(1,2))
plot(density(cardio$nFeature_RNA),main="Number of genes detected")
abline(v=500,col=2)
plot(density(cardio$nCount_RNA),main="Library size")
abline(v=2500,col=2)

cardio <- subset(cardio, subset = nFeature_RNA > 500 & nCount_RNA > 2500)
dim(cardio)
table(cardio$biorep)
```


# Run new integration with SCtransform normalisation

The new code does not work so I will take the same approach that I took integrating the samples within groups.

```{r}
cardio.list <- SplitObject(cardio, split.by = "biorep")
min(sapply(cardio.list, ncol))
```

```{r}
for (i in 1:length(cardio.list)) {
    cardio.list[[i]] <- SCTransform(cardio.list[[i]], verbose = FALSE)
}
```


```{r findanchors}
cardio.anchors <- FindIntegrationAnchors(object.list = cardio.list, dims=1:30,anchor.features = 3000,k.filter=216)
```

```{r integrate}
cardio.integrated <- IntegrateData(anchorset = cardio.anchors,dims=1:30)
```

# Perform clustering

```{r}
DefaultAssay(object = cardio.integrated) <- "integrated"
```

## Perform scaling and PCA

```{r pca}
cardio.integrated <- ScaleData(cardio.integrated, verbose = FALSE)
cardio.integrated <- RunPCA(cardio.integrated, npcs = 50, verbose = FALSE)
ElbowPlot(cardio.integrated,ndims=50)
```

```{r,fig.height=10, fig.width=12}
VizDimLoadings(cardio.integrated, dims = 1:4, reduction = "pca")
```

```{r}
DimPlot(cardio.integrated, reduction = "pca",group.by="orig.ident")
DimPlot(cardio.integrated, reduction = "pca",group.by="biorep")
DimPlot(cardio.integrated, reduction = "pca",group.by="sex")
DimPlot(cardio.integrated, reduction = "pca",group.by="batch")
```

```{r,fig.height=10, fig.width=12}
DimHeatmap(cardio.integrated, dims = 1:15, cells = 500, balanced = TRUE)
DimHeatmap(cardio.integrated, dims = 16:30, cells = 500, balanced = TRUE)
DimHeatmap(cardio.integrated, dims = 31:45, cells = 500, balanced = TRUE)
```

## Perform nearest neighbours clustering

```{r clust}
cardio.integrated <- FindNeighbors(cardio.integrated, dims = 1:20)
cardio.integrated <- FindClusters(cardio.integrated, resolution = 0.1)
table(Idents(cardio.integrated))
```

```{r}
par(mar=c(5,4,2,2))
barplot(table(Idents(cardio.integrated)),ylab="Number of cells",xlab="Clusters")
title("Number of cells in each cluster")
```


## Visualisation with TSNE

```{r tsne}
set.seed(10)
cardio.integrated <- RunTSNE(cardio.integrated, reduction = "pca", dims = 1:20)
```


```{r}
DimPlot(cardio.integrated, reduction = "tsne",label=TRUE,label.size = 6,pt.size = 0.5)+NoLegend()
```

```{r}
pdf(file="./output/Figures/NormalDev/tsne-cardioALL-res01.pdf",width=10,height=8,onefile = FALSE)
DimPlot(cardio.integrated, reduction = "tsne",label=TRUE,label.size = 6,pt.size = 0.5)+NoLegend()
dev.off()
```


```{r}
DimPlot(cardio.integrated, reduction = "tsne", group.by = "orig.ident")
```

```{r}
DimPlot(cardio.integrated, reduction = "tsne", split.by = "orig.ident")
```

```{r}
DimPlot(cardio.integrated, reduction = "tsne", group.by = "biorep")
```


```{r}
DimPlot(cardio.integrated, reduction = "tsne", group.by = "sex")
```

```{r}
DimPlot(cardio.integrated, reduction = "tsne", split.by = "sex")
```

```{r}
DimPlot(cardio.integrated, reduction = "tsne", group.by = "batch")
```


```{r,fig.width=12,fig.height=4}
par(mfrow=c(1,1))
par(mar=c(4,4,2,2))
tab <- table(Idents(cardio.integrated),cardio.integrated$biorep)
barplot(t(tab/rowSums(tab)),beside=TRUE,col=ggplotColors(9),legend=TRUE)
```

```{r,fig.width=12,fig.height=4}
par(mfrow=c(1,1))
par(mar=c(4,4,2,2))
tab <- table(Idents(cardio.integrated),cardio.integrated$orig.ident)
barplot(t(tab/rowSums(tab)),beside=TRUE,col=ggplotColors(3))
legend("topleft",legend=colnames(tab),fill=ggplotColors(3))
```

# Visualisation with clustree

```{r}
clusres <- c(0.1,0.2,0.3,0.4,0.5,0.6,0.7,0.8,0.9,1,1.1,1.2)
```
```{r clustree}
for(i in 1:length(clusres)){
  cardio.integrated <- FindClusters(cardio.integrated, 
                                   resolution = clusres[i])
}
```

```{r}
pct.male <- function(x) {mean(x=="m")}
pct.female <- function(x) {mean(x=="f")}
pct.fetal <- function(x) {mean(x=="fetal")}
pct.young <- function(x) {mean(x=="young")}
pct.adult <- function(x) {mean(x=="adult")}
```

```{r,fig.width=10,fig.height=10}
clustree(cardio.integrated, prefix = "integrated_snn_res.")
```

```{r,fig.width=10,fig.height=10}
clustree(cardio.integrated, prefix = "integrated_snn_res.",
         node_colour = "sex", node_colour_aggr = "pct.female",assay="RNA")
```

```{r,fig.width=10,fig.height=10}
clustree(cardio.integrated, prefix = "integrated_snn_res.",
         node_colour = "sex", node_colour_aggr = "pct.male",assay="RNA")
```

```{r,fig.width=10,fig.height=10}
clustree(cardio.integrated, prefix = "integrated_snn_res.",
         node_colour = "orig.ident", node_colour_aggr = "pct.fetal",assay="RNA")
```

```{r,fig.width=10,fig.height=10}
clustree(cardio.integrated, prefix = "integrated_snn_res.",
         node_colour = "orig.ident", node_colour_aggr = "pct.young",assay="RNA")
```

```{r,fig.width=10,fig.height=10}
clustree(cardio.integrated, prefix = "integrated_snn_res.",
         node_colour = "orig.ident", node_colour_aggr = "pct.adult",assay="RNA")
```


# Save Seurat object

```{r}
DefaultAssay(cardio.integrated) <- "RNA"
Idents(cardio.integrated) <- cardio.integrated$integrated_snn_res.0.1
```

```{r}
saveRDS(cardio.integrated,file="./output/RDataObjects/cardio-int-FYA-filtered.Rds")
#cardio.integrated <- readRDS(file="./output/RDataObjects/cardio-int-FYA.Rds")
# Load unfiltered counts matrix for every sample (object all)
load("./output/RDataObjects/all-counts.Rdata")
```

```{r}
columns(org.Hs.eg.db)
ann <- AnnotationDbi:::select(org.Hs.eg.db,keys=rownames(all),columns=c("SYMBOL","ENTREZID","ENSEMBL","GENENAME","CHR"),keytype = "SYMBOL")
m <- match(rownames(all),ann$SYMBOL)
ann <- ann[m,]
table(ann$SYMBOL==rownames(all))
```

```{r}
mito <- grep("mitochondrial",ann$GENENAME)
length(mito)
ribo <- grep("ribosomal",ann$GENENAME)
length(ribo)
missingEZID <- which(is.na(ann$ENTREZID))
length(missingEZID)
```

# Find Markers

```{r markers}
#adultmarkers <- FindAllMarkers(adult.integrated, only.pos = TRUE, min.pct = 0.25, logfc.threshold = 0.25)
# Limma-trend for DE
m <- match(colnames(cardio.integrated),colnames(all))
all.counts <- all[,m]
```

```{r}
chuck <- unique(c(mito,ribo,missingEZID))
length(chuck)
all.counts.keep <- all.counts[-chuck,]
ann.keep <- ann[-chuck,]
table(ann.keep$SYMBOL==rownames(all.counts.keep))
```

```{r}
numzero.genes <- rowSums(all.counts.keep==0)

#avg.exp <- rowMeans(cpm.DGEList(y.kid,log=TRUE))

#plot(avg.exp,numzero.genes,xlab="Average log-normalised-counts",ylab="Number zeroes per gene")

table(numzero.genes > (ncol(all.counts.keep)-20))

keep.genes <- numzero.genes < (ncol(all.counts.keep)-20)
table(keep.genes)

all.keep <- all.counts.keep[keep.genes,]
dim(all.keep)

ann.keep <- ann.keep[keep.genes,]
```


```{r}
y.cardio <- DGEList(all.keep)

logcounts <- normCounts(y.cardio,log=TRUE,prior.count=0.5)

#logcounts.n <- normalizeBetweenArrays(logcounts, method = "cyclicloess")

maxclust <- length(levels(Idents(cardio.integrated)))-1

grp <- paste("c",Idents(cardio.integrated),sep = "")
grp <- factor(grp,levels = paste("c",0:maxclust,sep=""))

design <- model.matrix(~0+grp+cardio.integrated$biorep)
colnames(design)[1:(maxclust+1)] <- levels(grp)

mycont <- matrix(0,ncol=length(levels(grp)),nrow=length(levels(grp)))
colnames(mycont)<-levels(grp)
diag(mycont)<-1
mycont[upper.tri(mycont)]<- -1/(length(levels(factor(grp)))-1)
mycont[lower.tri(mycont)]<- -1/(length(levels(factor(grp)))-1)

# Fill out remaining rows with 0s
zero.rows <- matrix(0,ncol=length(levels(grp)),nrow=(ncol(design)-length(levels(Idents(cardio.integrated)))))
test <- rbind(mycont,zero.rows)


fit <- lmFit(logcounts,design)
fit.cont <- contrasts.fit(fit,contrasts=test)
fit.cont <- eBayes(fit.cont,trend=TRUE,robust=TRUE)

fit.cont$genes <- ann.keep

summary(decideTests(fit.cont))

treat <- treat(fit.cont,lfc=0.5)

dt <- decideTests(treat)
summary(dt)

```

```{r,fig.width=10,fig.height=10}
par(mfrow=c(3,3))
for(i in 1:ncol(mycont)){
  plotMD(treat,coef=i,status = dt[,i],hl.cex=0.5)
  abline(h=0,col=colours()[c(226)])
  lines(lowess(treat$Amean,treat$coefficients[,i]),lwd=1.5,col=4)
}
```

## Write out marker genes for each cluster

```{r eval=FALSE}
contnames <- colnames(mycont)

for(i in 1:length(contnames)){
  topsig <- topTreat(treat,coef=i,n=Inf)
  write.csv(topsig,file=paste("./output/MarkerAnalysis/Cardiomyocytes/Development/Filtered/Cluster-",contnames[i],".csv",sep=""))
}
```

```{r}
fdr <- apply(treat$p.value, 2, function(x) p.adjust(x, method="BH"))
output <- data.frame(treat$genes,LogFC=treat$coefficients,AveExp=treat$Amean,tstat=treat$t, pvalue=treat$p.value, fdr=fdr)
write.csv(output,file="./output/MarkerAnalysis/Cardiomyocytes/Development/Filtered/MarkerAnalysis.csv")
```


# Perform gene set testing on C2 and GO sets

```{r}
contnames <- colnames(mycont)

load("./output/RDataObjects/human_c2_v5p2.rdata")
load("./output/RDataObjects/human_c5_v5p2.rdata")

c2.id <- ids2indices(Hs.c2,treat$genes$ENTREZID)
c5.id <- ids2indices(Hs.c5,treat$genes$ENTREZID)

reactome.id <-c2.id[grep("REACTOME",names(c2.id))]

c2.c0 <- cameraPR(treat$t[,1],c2.id)
reactome.c0 <- cameraPR(treat$t[,1],reactome.id)
go.c0 <- cameraPR(treat$t[,1],c5.id)

for(i in 1:length(contnames)){
  write.csv(cameraPR(treat$t[,i],c2.id),file=paste("./output/MarkerAnalysis/Cardiomyocytes/Development/Filtered/GeneSetTests/c2-",contnames[i],".csv",sep=""))
  write.csv(cameraPR(treat$t[,i],reactome.id),file=paste("./output/MarkerAnalysis/Cardiomyocytes/Development/Filtered/GeneSetTests/reactome-",contnames[i],".csv",sep=""))
  write.csv(cameraPR(treat$t[,i],c5.id),file=paste("./output/MarkerAnalysis/Cardiomyocytes/Development/Filtered/GeneSetTests/go-",contnames[i],".csv",sep=""))
}
```

# Check quality of clusters

The quality of the clusters look good.

```{r}
par(mfrow=c(1,1))
numgenes <- colSums(all.keep!=0)
boxplot(numgenes~grp)
```


# Heatmap of marker genes

```{r}
sam <- factor(cardio.integrated$biorep,levels=c("f1","f2","f3","y1","y2","y3","a1","a2","a3"))
newgrp <- paste(grp,sam,sep=".")
newgrp <- factor(newgrp,levels=paste(rep(levels(grp),each=9),levels(sam),sep="."))
o <-order(newgrp)
clust <- rep(levels(grp),each=9)
samps <- rep(levels(sam),length(levels(grp)))
```

## Summarise expression across cells 

```{r}
sumexpr <- matrix(NA,nrow=nrow(logcounts),ncol=length(levels(newgrp)))
rownames(sumexpr) <- rownames(logcounts)
colnames(sumexpr) <- levels(newgrp)

for(i in 1:nrow(sumexpr)){
  sumexpr[i,] <- tapply(logcounts[i,],newgrp,mean)
}

```

```{r sigheatmap, fig.width=12,fig.height=12}
sig.genes <- gene.label <- vector("list", length(levels(grp)))
for(i in 1:length(sig.genes)){
  top <- topTreat(treat,coef=i,n=Inf)
  sig.genes[[i]] <- rownames(top)[top$logFC>0][1:10]
  gene.label[[i]] <- paste(rownames(top)[top$logFC>0][1:10],levels(grp)[i],sep="-")
} 

csig <- unlist(sig.genes)
genes <- unlist(gene.label)

myColors <- list(Clust=NA,Sample=NA)
myColors$Clust<-sample(ggplotColors(length(levels(grp))),length(levels(grp)))
names(myColors$Clust)<-levels(grp)
myColors$Sample <- sample(ggplotColors(length(levels(sam))),length(levels(sam)))
names(myColors$Sample) <- levels(sam)


pdf(file="./output/Figures/NormalDev/cardio-heatmap-siggenes-summarised-FYA-filtered.pdf",width=20,height=20,onefile = FALSE)
aheatmap(sumexpr[csig,],Rowv = NA,Colv = NA, labRow = genes,
         annCol=list(Clust=clust,Sample=samps),
         annColors=myColors,
         fontsize=16,color="-RdYlBu",
         scale="none")
dev.off()
```

```{r, fig.width=10,fig.height=12}
aheatmap(sumexpr[csig,],Rowv = NA,Colv = NA, labRow = genes,
         annCol=list(Clust=clust,Sample=samps),
         annColors=myColors,
         fontsize=16,color="-RdYlBu",
         scale="none")
```

# Heatmap of pre-identified heart genes

```{r}
hm <- read.delim("./data/heart-markers-long.txt",stringsAsFactors = FALSE)
hgene <- toupper(hm$Gene)
hgene <- unique(hgene)

m <- match(hgene,rownames(sumexpr))
m <- m[!is.na(m)]

mycelltypes <- hm$Celltype[match(rownames(sumexpr)[m],toupper(hm$Gene))]
mycelltypes <- factor(mycelltypes)

mygenes <- rownames(sumexpr)[m]
mygenelab <- paste(mygenes,mycelltypes,sep="_")

pdf(file="./output/Figures/NormalDev/cardio-heatmap-hmarkers-summarised-FYA-filtered.pdf",width=20,height=15,onefile = FALSE)
aheatmap(sumexpr[m,],Rowv = NA,Colv = NA, labRow = mygenelab,
         annCol=list(Clust=clust,Sample=samps),
 #        annRow=list(Celltypes=mycelltypes),
         annColors=myColors,
         fontsize=14,color="-RdYlBu")
dev.off()

```

```{r, fig.width=10,fig.height=12}
aheatmap(sumexpr[m,],Rowv = NA,Colv = NA, labRow = mygenelab,
         annCol=list(Clust=clust,Sample=samps),
 #        annRow=list(Celltypes=mycelltypes),
         annColors=myColors,
         fontsize=14,color="-RdYlBu")
```





